Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance

### 2019-04-18 Manuscript/manuscript.Rmd
manuscript edits

Multiple random forest classification algorithms were built on domain data extracted from Hi-C data for 2 cell lines $C = \{GM12878, K562\}$, at four different resolutions $R = \{\text{10 kb}, \text{25 kb}, \text{50 kb}, \text{100 kb\}}$. For each classifier, we compared model performances using 3 separate predictor types $P = \{\text{OC}, \text{OP}, \text{Distance}\}$ and 4 class balancing techniques $B = \{None, RUS, ROS, SMOTE\}$. A system produced from the ensemble framework was defined as: $S = \{c, r, p, b\}$, where $c \in C$, $r \in R$, $p \in P$, and $b \in B$. Letting $|X|$ denote the cardinality of each set, we evaluated $|C| \times |R| \times |P| \times |B|$ different ensemble systems, totaling $96 (= 2 \times 4 \times 3 \times 4)$ separate models.

@@

We first evaluated performances for models implemented using OC predictor types, without re-sampling, as these characteristics represented models with the least granularity and could act as a baseline. In evaluating such composite metrics like F1-Score and MCC, we found that these models exhibited poor performances across resolutions, for both cell lines (Figure 5A). Furthermore, at some resolutions , the F1-Score and MCC could not be determined given by the model's complete inability to correctly classify any genomic bins that overlapped with flanked TAD boundaries in the test data, thereby resulting in 0 true positives <!--, as can be seen by the missing red points-->. When including OP perdictor types in the models, coupled with no resampling, we found that performances were worse than when using OC predictor types (Figure 5B). As with OC predictor types, these models were also not able to yield F1-Scores and MCCs at the 50 kb and 100 kb resolution for the GM12878 cell line or any resolution lower than 10 kb for the K562 cell line. Models built on distance-type predictors, however, exhibited slightly better performances than the other two types, yet still yielded missing or low F1-Scores and MCC at lower resolutions (Figure 5C). Models built on 100 kb resolutions were not able to yeild F1-Scores for either cell line. This indicated that the key driver leading to poor performances was not due to feature engineering, but instead the severity in the class imbalance.
<!--Mention AUC is not good metric using supplementary table-->

>>>>

Multiple random forest classification algorithms were built on domain data extracted from Hi-C data for 2 cell lines $C = \{\textit{GM12878}, \textit{K562}\}$, at four different resolutions $R = \{\textit{10 kb}, \textit{25 kb}, \textit{50 kb}, \textit{100 kb\}}$. For each classifier, we compared model performances using 3 separate predictor types $P = \{\textit{OC}, \textit{OP}, \textit{Distance}\}$ and 4 class balancing techniques $B = \{\textit{None}, \textit{RUS}, \textit{ROS}, \textit{SMOTE}\}$. A system produced from the ensemble framework was defined as: $S = \{\textit{c}, \textit{r}, \textit{p}, \textit{b}\}$, where $c \in C$, $r \in R$, $p \in P$, and $b \in B$. Letting $|X|$ denote the cardinality of each set, we evaluated $|C| \times |R| \times |P| \times |B|$ different ensemble systems, totaling $96 (= 2 \times 4 \times 3 \times 4)$ separate models.

@@

We first evaluated performances for models implemented using OC predictor types, without re-sampling, as these characteristics represented models with the least granularity and could act as a baseline. In evaluating such composite metrics like F1-Score and MCC, we found that these models exhibited poor performances across resolutions, for the GM12878 cell line (Figure 5A). The average F1-Score and MCC across resolution was found to be 0.04 and 0.07 respectively. When including OP perdictor types as the feature space, coupled with no resampling, we found that performances were worse than when using OC predictor types for the GM12878 cell line (Figure 5B). Here, the average F1-Score and MCC was found to be 0.01 and 0.03 respectively. Models built on distance-type predictors in conjuction with no resampling, however, were found to exhibit slightly better performances than the other two types, with average F1-Scores and MCCs of 0.08 and 0.10 respectively (Figure 5C). Importantly, when validating models built on fully imbalanced datasets, the F1-Score and MCC could not be determined at lower resolutions given by the model's complete inability to correctly classify any genomic bins that overlapped with flanked TAD boundaries in the test data, thereby resulting in 0 true positives. Furthermore, evaluating models using conventional metrics such as AUC or accuracy did not depict the true severity in performance. The average AUC across resolutions for the OC, OP, and distance type predictors was 0.65, 0.66, and 0.85 respectively, where as the average accuracies were 0.89, 0.89, and 0.90 respectively (Supplementary Table 3).
